### 2019-09-15 ic-journal/response-letter.tex

\newcommand{\Mark}{{\underline{\tt R:} }}

\begin{center}
\textbf{
{\small Changes for Submission 1 to SI:EXPRESS/SOS'18}\\[5mm]
{\large Unique Solutions of Contractions, CCS, and their HOL
  Formalisation
}}
\end{center}

\vskip 10pt
\noindent

We would like to thank the reviewers for many useful
comments and suggestions. We believe we have taken all of them into consideration, as best
as we could.

The main addition, here again following the recommendations of the reviewers 
has been the extension of the our central theorems 
to the multivariable case (ie., moving from ``unique solution theorems'' for a single
equations, or contractions,  to multi-equations  or multi-contractions). 
 This has been a significant effort (also because a few alternative ways to make the
 extension had to be examined). At the end, this work 
consists of another 4,000 lines of proof scripts and a whole new
Section 5 in the paper.

In the following, we discuss the changes to the paper and provide a
point-by-point reply to the issues that have been raised.

Our answers are marked  with an `` \Mark''  at the beginning.

\finish{add some kind of mark, like above, so that our answers are clearly seen} 
\ \\

@@

\item \begin{quote}
    In general, these kinds of results are error-prone, especially
    those related to weak bisimilarity: there are multiple known
    mistakes in the literature. Therefore, formalisation is a
    worthwile effort. Moreover, the formalisation itself is quite
    readable. The paper itself documents this formalisation and is,
    for the most part, clear.
  \end{quote}

\item \begin{quote}
    The paper is an extended version of an EXPRESS/SOS paper. The
    extension consists of a more detailed report of the proofs -
    perhaps this is enough.
  \end{quote}

\item \begin{quote}
    I was disappointed though that multi-variable contexts are still
    not treated (I reviewed the EXPRESS/SOS paper as well and
    complained about that then). This means the current formalisation
    does *not* fully formalise the unique-solution method. The journal
    version would have been ideal to solve this issue. It is briefly
    discussed in the conclusion, but in fact, the first author's
    master thesis argues that it is possible: "With these new devices,
    it’s possible to formalize all “unique solution of equations”
    theorems with multi-variable equations, without touching existing
    formalization framework.". The current discussion should at the
    very least explain this (is the master thesis wrong? is this too
    hard, or too much work?), but hopefully it can just be done
    properly so that the paper fully covers the correctness of the
    unique-solution method. 
  \end{quote}

\Mark  As indicated  above,  we have make the extension to the multivariate (i.e. multi-variable)
  versions of two core theorems:  the \emph{unique solution of rooted contractions theorem}
  (Theorem~\ref{t:rcontraBisimulationU}) and \emph{Milner's strong
  unique-solution theorem} (Prop.~4.14 (2)
  of~\citep[p.~103]{Mil89}), and added a whole new section to  discuss the main issues
  faced in the  extension (having to do, eg, with 
  multivariable  contexts, weakly-guarded equations and multivariable substitutions). 
The extension  is indeed based on the existing framework in
  the sense that, we do not  change the CCS datatype definition
  and the  SOS rules

@@

\item \begin{quote}
    L91 "bisimiarity"
  \end{quote}

Fixed.

\item \begin{quote}
    proof of thm 3.10: could there be some kind of picture to help? also, the "Hence," in the one-to-last line refers only to the property right before it, right? (about $C[\tilde{Q}]$)
  \end{quote}

  We have rephrased this proof (also the theorem statements) with a
  picture (Fig.~\ref{fig:310} illustrating the proof. 
% The "Hence" part
%   actually used a property of $\wb$ that can be easily proven by induction
%   on each $\tau$-transition in a weak transition.

@@

\item \begin{quote}
    L209: "their precongruences" precongruence closure?
  \end{quote}

  Rephrased.

% : ``... because all we need from $\mcontrBIS$ (or
% $\rcontr$) is their substitutivity properties (as precongruences)
% w.r.t the corresponding definitions of contexts.''

@@

\item \begin{quote}
    L224: full stop and new sentence after "1972"
  \end{quote}
  Done.

\item \begin{quote}
    L227 "considerable" -> "considerably"
  \end{quote}
  Done.

\item \begin{quote}
    L228 "formal theories built in HOL is easily convincible" questionable grammar
  \end{quote}

We have rephrased the sentence.
%   Now we used the following sentence: ``As a consequence, theories and proofs verified in HOL are relatively more
%   convincible to non-experts.''
\item \begin{quote}
    L238 how is this without loss of generality?
  \end{quote}

  Sentence eliminated as now we have formalized the multivariate
  version of the unique-solution theorems. The ``without loss of
  generality'' intended to point out  that all steps of
  informal proofs do appear in the univariate formal proofs. The
  additional difficulities for the formal multivariate proof are discussed in Section~\ref{sec:multivariate}.

@@

\item \begin{quote}
    L245 maybe more clear to just give the definition of $\beta$ Action 
  \end{quote}

  The original HOL definition of the type Action is now given.

\item \begin{quote}
    L250,300: perhaps explain what these (double) backquotes signify or just remove the sentence
  \end{quote}
  Sentences removed. In HOL, single and double backquotes are
  delimitator between ``inner'' (logic) and ``outer'' (proof)
  languages but this is not really  relevant for the present paper.
  After new HOL releases (k13) most uses of backquotes in  definitions can be eliminated.

@@

\item \begin{quote}
    L280 (around): put some of these equalities on a single line? enough space
  \end{quote}
  Now we have removed most less-interesting cases and explained in
  details the two interesting cases (recursion operator and (free) variables).

\item \begin{quote}
    L316 and later, e.g. 323: is "p" a closed ccs term?
  \end{quote}

\item \begin{quote}
    L361 "Many lemmas ... contexts" this is a bit vague
  \end{quote}

We have rephrased
% : ``Many lemmas about above concepts (\texttt{CONTEXT}, \texttt{WG},
%   \texttt{SEQ}, etc.)''

\item \begin{quote}
    L374 "An highlight"
  \end{quote}
  Fixed.  %Now we say ``One highlight...''.

\item \begin{quote}
    L378 "is a" => "are"
  \end{quote}
  Fixed.  
%``... what kind of binary relations on CCS process are (strong) bisimulations.''

@@

\item \begin{quote}
    L396 "we call *the* Hol_coreln"
  \end{quote}
  Fixed.

\item \begin{quote}
    L410 $E$ and $E'$ missing
  \end{quote}
  All formal theorems and definitions are generated from HOL4.
  By default, all theorems are fully specialised, removing outermost
  universal quantifiers. For certain important theorems like this one,
  now we have disabled the specialization showing outermost
  universal quantifiers.

\item \begin{quote}
    L436 "is defined upon weak transition" grammar
  \end{quote}

We have rephrased: ``The definition of weak bisimulation is based on weak
  and $\epsilon$--transitions:''

\item \begin{quote}
    L505 this is good: it is missing for strong bisimilarity, should
    be added there too (the two definitions are now not explicitly
    related)
  \end{quote}
  We have removed the duplicated part and only kept the part for weak bisimulation.

\item \begin{quote}
    L580 "a contraction" ?
  \end{quote}
We have rephrased this

@@

\item \begin{quote}
    section 4.6 header: "the formalisation of bisimulation up to
    bisimilarity", i would say; also since you're not using
    compatible/respectful functions
  \end{quote}
  Fixed.

\item \begin{quote}
    L599 PRQ not so pretty spacing. perhaps mention around here that
    naive weak bisimulation up to weak bisimilarity is unsound?
  \end{quote}
  Fixed ( we  mention that the ``naive weak bisimulation up to weak
  bisimilarity is unsound'' right before the definition of ``weak
  bisimilation up to.'')

\item \begin{quote}
    L655 (around here): i'd say in the text above it that this is a
    different bisimulation game, not a different up-to technique
  \end{quote}
  Thanks. We have rephrased it 
%added ``we actually used a different bisimulation game defined below''.
\item \begin{quote}
    L708 " closure of bisimilarity under such operator" not sure whether this is gramatically correct
  \end{quote}
  Rephrased
% : `` we also consider the relation closure of bisimilarity
%   w.r.t. summation...''

\item \begin{quote}
    figure 2: I'm a bit confused since the other implication is missing. 
  \end{quote}
  We fixed one minor issue in Fig.~2 (from \texttt{SUM_EQUIV} to
  $\rapprox$ the label $\subseteq$ should be $\supseteq$). The main
  message of Fig.~2 is to show the equivalences of three involved
  relations by making a loop of implications.

@@

\item \begin{quote}
    L740 "this proof" which? 
  \end{quote}

  ``this proof (Theorem~\ref{thm:coarsestfiniteState} in next
  section)``.

\item \begin{quote}
    The beginning of section 4.8 is really confusing: what are "the
    two root processes"? What are we going to prove in this section,
    and why, what is the aim here? Again, under which assumptions are
    we going to work?
  \end{quote}

  We have re-organized this section, putting the target theorem and
  its limitations  first.

@@

\item \begin{quote}
    L820 "bisimularity"
  \end{quote}
  Fixed.

\item \begin{quote}
    Lemma 4.14 is nicely introduced, Thm 4.15 not. 
  \end{quote}

  We have add some explanation on the proof details of Thm 4.15.

@@

  We have removed the one before Example 3.7.

\item \begin{quote}
    The notion of “substitutivity property” should be introduced, now it’s used but not explained.
  \end{quote}

\item \begin{quote}
    I strongly believe that Section 4 is accessible only to people
    already familiar with HOL syntax. I suggest the authors to give
    some more details. Some examples.
  \end{quote}

  We have added some HOL background and links to further details about its official documentation.

@@

\item \begin{quote}
    You assume the reader knows what Hol_reln is, but this not always the case. 
Then, you define the congruence in terms of the equivalence, but is
the equivalence already defined somewhere?
  \end{quote}

  We have  further explained what  \texttt{Hol_reln} does and returns.

@@

\item \begin{quote}
    The argument at lines 387-389 is clear only after the lines 394-427 have been read.
  \end{quote}

  We have merged the strong and weak cases.

\item \begin{quote}
    At lines 577 you write that proving properties for contraction is
    in general harder than proving properties for expansion. Could you
    explain why?
  \end{quote}

  We added some explanations, including this sentence:
 ``this is mostly due to the (surprising) fact
 that, contraction does not imply weak bisimulation (while expansion
 indeed implies weak bisimulation) ''.

@@

\item \begin{quote}
    L32: Add the word 'also' after 'the work is'
  \end{quote}
  Done.

\item \begin{quote}
    L34: From the view of -> From the point of view of
  \end{quote}
  Done.

\item \begin{quote}
    L35: Regarding the sentence "as formally proving a previously known result gives us the chance to see
    what is really needed for establishing that result."
    Either omit this sentence or explain what technical insights one
    gains by doing this work
  \end{quote}

  Our formal proof of the standard  result  ``$\rapprox$ is the
  coarsest congruence contained in $\wb$'' is one such example, where
  the formal theorem we obtained has slightly weaker antecedents under
  the same proof idea. We have however now omitted the sentence.

@@

\item \begin{quote}
    L55-60: Remove "Given a deadlock 0"
  \end{quote}
  Done.

\item \begin{quote}
    L58: the trailing 0 -> a trailing 0
  \end{quote}
  Done.

\item \begin{quote}
    L59: "shown in Fig. 1" is placed after "a Labeled Transition
    System", causing the belief that Figure 1 contains an LTS. 
    But there is no LTS. Figure 1 contains the SOS of CCS.
  \end{quote}

  We have rephrased it (pointing out that Fig.1 represents an LTS expressed in SOS rules).

@@

\item \begin{quote}
    Semantics: The relabeling rule is not explained anywhere.
  \end{quote}

  In this project, relabeling is not discussed besides their presense
  in CCS syntax. In particular, relabeling doesn't bring in any delicate issues
  in the proofs of unique-solution theorems.  We have added a footnote to clarify this.

\finish{is the footnote still there? anyway, i do not like footnotes. One could just say
  two words on the rule for relabelling (maybe we do say it already?)}

\item \begin{quote}
    Semantics: The constraint on the relabeling functions does not belong to the inference rule, but should be formulated as a restriction on the syntax.
  \end{quote}
  We have moved the side condition to the main text where we explain
  the syntax.

\item \begin{quote}
    L74: Add the word 'a' between 'are' and 'context'.
  \end{quote}
  Done.

@@

\item \begin{quote}
  Contexts: Does an $i$-holed context have exactly one hole $[]_i$ for
  each $i$? This question is not answered anywhere in the text.

  In the formalisation you only use contexts with one kind of hole $[]$. This definition can be changed to only
  use one hole as well, with a clear indication that this hole can occur multiple times.
\end{quote}

We have completely rewritten this part, now we adopt free variables
and unified CCS expressions for both processes and equations.
We have then explained the meaning of context in relation to this 
(having multiequations now we have to deal also with contexts with multiple holes)

@@

\item \begin{quote}
    L82: Remove the comma after $Q'$
  \end{quote}
  Done.

\item \begin{quote}
    Definition 2.2: Superscript $c$ is already used for contextual closure.
    Does $\approx^c$ stand for the contextual closure of $\approx$, or is it something else?
    Since contextual closure is not used anywhere in the paper, it is best to remove it.
    Otherwise you need different notation for rooted bisimilarity.
  \end{quote}
  We have removed the definition of ``context closure'' since it's not
  (explicitly) used in the paper.  Then $\rapprox$ is a standard
  notation for rooted bisimilarity (eg  Gorrieri~\cite{Gorrieri:2015jt})

\item \begin{quote}
    L91: What does 'reducing' mean?
  \end{quote}
We have rephrased the sentence.

\finish{above: check, i think i rewrote it}

@@

\item \begin{quote}
    L92: Add a colon after 'results'
  \end{quote}

\item \begin{quote}
    L118: Why is the word 'actions' in bold font? It says 'ds' in the margin, what does that mean?
  \end{quote}
  We have corrected it.

\item \begin{quote}
Definition 3.3: The definition of ‘sequential’ is awkward.  Moreover
this definition calls for some examples since the definitions can be
very confusing.
\end{quote}
We have added some examples, and used Milner's original definition of sequentiality.

\item \begin{quote}
Theorem 3.4: What are direct sums?
\end{quote}
Rephrased: ``with guarded sums only''. We have removed all usages of
``direct sum''.

\item \begin{quote}
    L.127: I think it would be clearer to give two concrete solutions
    rather than saying that any process not using a is a solution.
  \end{quote}
  Done. (Two examples are $\nil$ and $b.\nil$.)

@@

Corrected.

\item \begin{quote}
    L184: closes -> completes
  \end{quote}
  Done.

\item \begin{quote}
    L194: ‘the definition should not be recursive’?
  \end{quote}

  We meant the definition should not be \emph{coinductive}, just like the
  case of $\rapprox$. Notice that, this is just an informal idea when searching for
  a possible definiton of rooted contraction, and the definition we
  found at last is indeed not coinductive

\item \begin{quote}
    Definition 3.11: Similar to the remark about Definition 2.2.
  \end{quote}

We have tried to clarify this.
  % You meant removing also the comma (period) here? But it's different
  % here: the definition ends at clause (2), an end period is added with
  % an ``enspace'' distance.

\item \begin{quote}
    Proof of Theorem 3.14: Why do you use context $C$ instead of tuples of expressions $\tilde{E}$?
    If this is the case because you only use contexts with one kind of hole, then you should use the letter
    E instead of C.
    If $\tilde{P}$ is a solution then $P_i >=^c E_i[\tilde{P}]$ for all i, but this quantification over i is missing.
  \end{quote}
  We have  re-worked this proof.

@@

\item \begin{quote}
    L228: what does convincible mean here?
  \end{quote}
  We have further explained it.

\item \begin{quote}
    L230: single?
  \end{quote}
  Changed to: ``the \emph{same} (programming language)''

\item \begin{quote}
    L235: The sentence between parentheses is an important remark and
    should not be between parentheses.
  \end{quote}
  Removed parentheses.

\item \begin{quote}
    L238: without loss of generality?
  \end{quote}
  We have eliminated this sentence as now we have  the
  multivariate formalizations.

\item \begin{quote}
    L239: similar to the comment about line 118, theorem-prover is
    bold and 'ds' is in the margin.
  \end{quote}
  We have eliminated this sentence as now we have  the
  multivariate formalizations.

@@

\item \begin{quote}
    L245: now based on HOL’s option theory?
  \end{quote}
  \texttt{option} is the name of the HOL formal theory.

\item \begin{quote}
    L249: What are ‘backquotes’?
  \end{quote}
  We have eliminated most uses of backquotes using HOL's new syntax.

\item \begin{quote}
    Footnote 3: What are CCS literals?
  \end{quote}
Corrected.

\item \begin{quote}
    L265: What does the turnstile mean? I would expect that structural
    operational rules are axioms of the theory, but the turnstile
    suggests to me that they are theorems.
  \end{quote}
  We have tried to  clarify this. 
(SOS rules are  not axioms, because it's
  generally not allowed to define new axioms. SOS rules are actually
 theorems added by the inductive relation \texttt{TRANS} as a
  conservative extension of the logic.)

@@

\item \begin{quote}
    General: in the lines 293-300 the trailing deadlock 0 is given,
    while in the preliminaries you stated that it would be omitted. If
    you decide to keep the deadlock, it should also be in bold font.
  \end{quote}
  We have made them all in bold fonts.

\item \begin{quote}
    L296: I don’t understand fact (1).
  \end{quote}
We have re-phrased this part: ``(1) there exist indeed three
  transitions mentioned above; (2) there exists no other
  transition.''

\item \begin{quote}
    L299: ‘Hence ... CCS_TRANS_CONV’ is not a grammatically correct sentence; there is a verb missing.
  \end{quote}
  We have rephrased this sentence.

\item \begin{quote}
    L316: Explain what is p.
  \end{quote}
  We have explained it in the paper. In this case there is no hole in the context.

\item \begin{quote}
    Footnote 4: Do you mean ‘unguarded recursion’? How can relabelling operators cause infinite branching?
  \end{quote}
  We have removed the part ``due to the use of ...''. 
% Infinite
%   branching is indeed possible, no matter what's the reason.

@@

\item \begin{quote}
    L378: relations on CCS process -> relation on CCS processes
  \end{quote}
  Done.

\item \begin{quote}
    Footnote 5: Add the word 'respectively' between 'quantifiers' and 'in'.
  \end{quote}
  Done.

### 2019-09-15 ic-journal/response-letter.tex

\item \begin{quote}
    Footnote 6: Regarding the last sentence between parentheses. What do you mean by this?
  \end{quote}
  By default, all theorems (their \TeX code) exported from HOL are
  fully specialised, removing outermost universal quantifiers. We have
  moved this sentence to an earlier footnote on the first such exported theorem.

\finish{check above, after my modifications}

\item \begin{quote}
    General: The same trick of applying the coinduction package to the
    definition of bisimulation is repeated twice.
  \end{quote}
  We have removed the one for strong bisimulation, leaving
  only the part for weak bisimulation.

\item \begin{quote}
    General: In the sections about the HOL formalisation you refer to
    weak bisimulation instead of bisimulation.    
  \end{quote}
  In the first part of the paper, strong bisimulation rarely appears
  as the focus is on weak bisimulation. But in the formalisation
  sections we also formalised Milner's unique-solution theorem for
  strong bisimilarity and many statements must distinguish between
  strong/weak bisimilarities. Thus we often use ``weak bisimulation''
  to make it  clearer.

\item \begin{quote}
    L461: I think there is no need to include those 3
    theorems. Wouldn’t it suffice to write that similar theorems are
    returned as for \texttt{STRONG\_EQUIV}?
  \end{quote}
As pointed out above, we have  only maintained 
 the part for weak bisimulation.

\item \begin{quote}
    L496-503: A similar remark can be made for \texttt{STRONG\_EQUIV}. Why do you put it here?
  \end{quote}
  Same as above.

@@

\item \begin{quote}
    L523,524: Similar to the comment about line L28
  \end{quote}
  Done.

\item \begin{quote}
    L545: What does this statement mean?
  \end{quote}
  We have rephrased this sentence. (Basically we meant that
any  algebraic law for strong bisimilarity can be stated also for
  weak and rooted bisimilarity.)

@@

\item \begin{quote}
    Up-to techniques: You mention four up-to techniques for weak
    bisimilarity and each of them are not powerful enough to prove the
    unique solution theorems. Why are they not powerful enough?
  \end{quote}

  We have repharsed this part.
% , moving the citation
%   \cite{sangiorgi1992problem} (which explains the reason) right before
%   the first sentence.

\item \begin{quote}
    L580: ‘Surprisingly’ To me this is not really surprising. Why do you find it surprising?
  \end{quote}
 We have rephrased this sentence:
%   ``this is mostly due to the (surprising) fact
%   that, although the contraction preorder is contained in weak
% bisimilarity, contraction does not imply weak bisimulation (while
% expansion does so)''.

@@

\item \begin{quote}
    L723-L727: This sentence should be split into multiple sentences.
  \end{quote}
  We have rephrased the sentence.

\item \begin{quote}
    L744-745: pure mathematics?
  \end{quote}
  We have removed this sentence.

\item \begin{quote}
    L746: Arbitrary -> Arbitrarily
  \end{quote}
  Done.

\item \begin{quote}
    L749: "infinite such sub-processes" this construction does not seem right and could be avoided.
  \end{quote}
  We have rephrased the sentence: 
% ``There could be infinitely many such sub-processes, even on finitely
%   branching processes.''

### 2019-09-15 ic-journal/ccs.tex
text
Literally handled ALL review issues. DONE.
Final fixes; disabled highlights

\hlD{Here we recall the core theory of CCS.}

@@

\ldots\}$ \hlD{yielding} input and $\overline{\mbox{output}}$ labels (actions), plus a special invisible

@@

defined from $\nil$ \hl{(deadlock) and agent variables} by the operators

@@

\hlD{In our presentation of CCS, 

@@

We sometimes omit \hlD{a} trailing $\nil$, e.g., writing $a|b$ for $a.\nil |b .\nil$.
\hl{A CCS process $P$ may \hlD{evolve} to another one, say $P'$, under
  an action $\mu$, written by $P \arr{\mu} P'$.}
The \hlD{transition semantics} of CCS processes is given by means of

@@

% \paragraph{CCS contexts and equations}
% And $\ct{\til P}$ is the process obtained by replacing each hole
% $\holei i$ of the  context $\qct$ with $P_i$.  
% We write $\ctx \R$ for the closure of a relation under contexts.
% Thus $P\: \ctx \R\: Q$
% means that there \hl{exists a} context $\qct$ and tuples $\til P,\til Q$ with
% $P =  \ct{\til P}, Q =  \ct{\til Q}$ and $\til P \RR \til Q$.

>>>>

\hl{Here we recall the core theory of CCS.}

@@

\ldots\}$ \hl{yielding} input and $\overline{\mbox{output}}$ labels (actions), plus a special invisible

@@

defined from $\nil$ \hl{(the terminated process)} and agent variables by the operators

@@

\hl{In our presentation of CCS, 

@@

We sometimes omit \hl{a} trailing $\nil$, e.g., writing $a|b$ for $a.\nil |b .\nil$.
A CCS process $P$ may \hl{evolve} to another one, say $P'$, under
  an action $\mu$, written by $P \arr{\mu} P'$.
The \hl{transition semantics} of CCS processes is given by means of

@@

### 2019-09-15 ic-journal/coarsest.tex
Final fixes; disabled highlights

To \hl{actually} construct this process $k$,
 the proof relies on arbitrary infinite sums of 
processes and uses transfinite induction to obtain
an arbitrary large sequence of processes (firstly introduced by Jan
Willem Klop \cite{van2005characterisation})
 that are all pairwise non-bisimilar.
We have only partially formalised
this proof (Theorem~\ref{thm:coarsestfiniteState} in next section), because ``the typed logic
implemented in various HOL systems (including Isabelle/HOL) is not
strong enough to define a type for all possible
ordinals''~\cite{norrish2013ordinals}.
\hlD{As a consequence, the final result is only for finite-state
  CCS~--- a  restricted class.}

The assumption in the previous \texttt{PROP3_COMMON} requires the
existence of a special CCS process, which is not weakly bisimilar to
any sub-process emanating from the two root processes by weak
transitions.
\hl{There could be infinitely many such sub-processes, even on finitely
branching processes.}
We can, however, consider the equivalence classes of CCS processes
modulo weak bisimilarity.
If there are infinitely many of such classes, 
then it will be 
possible to choose one  that is distinct from all the (finitely many) states in the
graphs of the two given processes.  
This can be done following Klop's contruction \cite{van2005characterisation}.
 We call  ``Klop processes''  the processes in this construction:
\begin{definition}[Klop processes]
For each ordinal $\lambda$, and an arbitrary chosen non-$\tau$ action $a$,
define a CCS process $k_\lambda$ as follows:
\begin{itemize}
\item $k_0 = 0$,
\item $k_{\lambda+1} = k_\lambda + a.k_\lambda$ and
\item for $\lambda$ a limit ordinal, $k_\lambda = \sum_{\mu < \lambda}
  k_\mu$, meaning that $k_\lambda$ is constructed from all graphs
  $k_\mu$ for $\mu < \lambda$ by identifying their root.
\end{itemize}
\end{definition}
When processes are finite-state, that is,
the number of  states in which a process may evolve by performing
transitions is finite, 
we can use  the following subset of Klop processes, 
defined as a recursive function (on natural numbers) in HOL4:
\begin{definition}{(Klop processes as recursive function on natural numbers)}
\begin{alltt}
\HOLConst{KLOP} \HOLFreeVar{a} \HOLNumLit{0} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{\ensuremath{\mathbf{0}}}
\HOLConst{KLOP} \HOLFreeVar{a} \ensuremath{(}\HOLConst{SUC} \HOLFreeVar{n}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n} \HOLSymConst{\ensuremath{+}} \HOLConst{label} \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n}\hfill{[KLOP_def]}
\end{alltt}
\end{definition}

>>>>

To \hl{actually} construct this process $k$,
 the proof relies on arbitrary infinite sums of 
processes and uses transfinite induction to obtain
an arbitrary large sequence of processes (firstly introduced by Jan
Willem Klop \cite{van2005characterisation})
 that are all pairwise non-bisimilar.
We have only partially formalised
this proof (Theorem~\ref{thm:coarsestfiniteState} in next section), because ``the typed logic
implemented in various HOL systems (including Isabelle/HOL) is not
strong enough to define a type for all possible
ordinals''~\cite{norrish2013ordinals}.
As a consequence, the final result is only for finite-state
  CCS~--- a  restricted class.

The assumption in the previous \texttt{PROP3_COMMON} requires the
existence of a special CCS process, which is not weakly bisimilar to
any sub-process emanating from the two root processes by weak
transitions.
\hl{There could be infinitely many such sub-processes, even on finitely
branching processes.}
We can, however, consider the equivalence classes of CCS processes
modulo weak bisimilarity.
If there are infinitely many of such classes, 
then it will be 
possible to choose one  that is distinct from all the (finitely many) states in the
graphs of the two given processes.  
This can be done following Klop's contruction \cite{van2005characterisation}.
 We call  ``Klop processes''  the processes in this construction:
\begin{definition}[Klop processes]
For each ordinal $\lambda$, and an arbitrary chosen action $a \neq \tau$,
define a CCS process $k_\lambda$ as follows:
\begin{itemize}
\item $k_0 = 0$,
\item $k_{\lambda+1} = k_\lambda + a.k_\lambda$ and
\item for $\lambda$ a limit ordinal, $k_\lambda = \sum_{\mu < \lambda}
  k_\mu$, meaning that $k_\lambda$ is constructed from all graphs
  $k_\mu$ for $\mu < \lambda$ by identifying their root.
\end{itemize}
\end{definition}
When processes are finite-state, that is,
the number of  states in which a process may evolve by performing
transitions is finite, 
we can use  the following subset of Klop processes, 
defined as a recursive function (on natural numbers) in HOL4:
\begin{definition}{(Klop processes as recursive function on natural numbers)}
\begin{alltt}
\HOLConst{KLOP} \HOLFreeVar{a} \HOLNumLit{0} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{\ensuremath{\mathbf{0}}}
\HOLConst{KLOP} \HOLFreeVar{a} \ensuremath{(}\HOLConst{SUC} \HOLFreeVar{n}\ensuremath{)} \HOLSymConst{\HOLTokenDefEquality{}} \HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n} \HOLSymConst{\ensuremath{+}} \HOLConst{label} \HOLFreeVar{a}\HOLSymConst{\ensuremath{\ldotp}}\HOLConst{KLOP} \HOLFreeVar{a} \HOLFreeVar{n}\hfill{[KLOP_def]}
\end{alltt}
\end{definition}
